### 2020-07-14 paper.Rmd
some notes on validation

- Do simulations to show that k-variance anticlustering outperforms 
other objectives with regard to similarity in means / variances

- New objective seems to work best when N is high relative to K; 
optimizing two objectives at the same time is hard when N is low. Easier 
when only optimizing one objective (i.e., the mean, as done by normal
k-means anticlustering).

- Two examples: $K = 2$, $N = 20$ vs. $N = 100$, using classical
k-means: only mean; using k-variance: only variance; combination:
both, but works better for larger $N$

- Often, maximizing k-means-variance yields more similar variances 
than k-variance alone. Why is that? I guess: structure in the solution 
space, local maxima are real. When also considering means, you more 
often escape bad local maxima?

@@

- use some examples (e.g., OASIS data set)

# Test equivalence of objectives

Check that I can actually just append the squared difference data as 
additional columns to the original data frame:

squared_from_mean <- function(data) {
  apply(data, 2, function(x) (x - mean(x))^2)
}

N <- 500
M <- 20
data <- matrix(rnorm(N * M), ncol = M)
var_data <- squared_from_mean(data)

K <- 10
clusters <- sample(rep_len(1:K, N))

obj1 <- variance_objective(data, clusters) + variance_objective(var_data, clusters)
obj2 <- variance_objective(cbind(data, var_data), clusters)

tolerance <- 1e-10

obj1 - obj2 < tolerance

obj1
obj2

>>>>

- Simulation 1: Show that k-means extended obtains 99.999% of the 
k-means and k-variance objective, on average. 
  + New objective seems to work best when N is high relative to K; optimizing 
two objectives at the same time is hard when N is low. Easier when only 
optimizing one objective (i.e., the mean, as done by normal k-means 
anticlustering). What I can say: When N / K > X, the objective is 
usually 99.99% of both objectives.
  + Two examples: $K = 2$, $N = 20$ vs. $N = 100$, using classical
k-means: only mean; using k-variance: only variance; combination:
both, but works better for larger $N$

- Simulation 2: show that k-variance anticlustering outperforms 
other objectives with regard to similarity in means / variances 
(in particular: anticluster editing)

- Often, maximizing k-means-variance yields more similar variances 
than k-variance alone. Why is that? I guess: structure in the solution 
space, local maxima are real. When also considering means, you more 
often escape bad local maxima? The finding is: k-means-extended better 
minimizes the maximum difference between the variance across groups 
than k-variance (for K > 2), which in the eye of the researcher, may 
be a more intuitive criterion.

@@

- use some examples (e.g., OASIS data set, 900 items allocated to 10 
sets, same mean and variance across the 10 sets)
